---
title: "HW7"
author: "Enguang Fan"
date: "3/19/2020"
output:
  pdf_document: default
  html_document: default
---



## City crime data
0. Inspect the help file of the **all.us.city.crime.1970** data in the **cluster.datasets** package to familiarize yourself with it. We will be clustering cities based on the crime rate variables in columns 5-10.

```{r}
library(cluster.datasets)
set.seed(5)
data(all.us.city.crime.1970)
help(all.us.city.crime.1970)
#df <- data(all.us.city.crime.1970)
```


1. For K=2,3,4,...,10 find the within cluster sum of squares and plot this against K. Based on this, what would be a reasonable number of clusters?

```{r}
## save WSS(K) for K=2,10)
WSS=NULL
for(K in 2:10){
 kms=kmeans(all.us.city.crime.1970[,5:10],centers=K,nstart=10)
 wssk=sum(kms$withinss)
 WSS=c(WSS,wssk)
}

plot(c(2:10),WSS,xlab="K",ylab="WSS(K)")
```
There is an elbow when K = 4 or 5, so I'd use K = 4


2. Find the K-means solution for the number K you found in 1. Which cities go in which clusters?

```{r}
set.seed(5)
cityclusters=kmeans(all.us.city.crime.1970[,5:10],centers=4)
city = all.us.city.crime.1970[,1]
```

```{r}
print("cities in cluster 1")
city[cityclusters$cluster==1]
```

```{r}
print("cities in cluster 2")
city[cityclusters$cluster==2]
```

```{r}
print("cities in cluster 3")
city[cityclusters$cluster==3]
```

```{r}
print("cities in cluster 4")
city[cityclusters$cluster==4]
```




3. Now find K based on maximizing the average silhouette. What is the optimal K based on this criterion?

```{r}
set.seed(5)
#use the function from lecture notes
avsil=function(x,kmobject){
## x is the raw data
## kmobject is the resulting list after running kmeans
 n=nrow(x)
 K=length(kmobject$size)
 sils=c(1:n)
for(i in 1:n){
 distvals=c(1:K)
for(k in 1:K){
 distvals[k]=sum((x[i,]-kmobject$centers[k,])^2)
 }
 distvals=sort(distvals)
 sils[i]=(distvals[2]-distvals[1])/distvals[2]
 }
return(mean(sils))
}

AS=c(1:9)
for(k in 2:10){
 km=kmeans(all.us.city.crime.1970[,5:10],centers=k,nstart=10)
 AS[k-1]=avsil(all.us.city.crime.1970[,5:10],km)
}
plot(c(2:10),AS)

```

The optimal K = 5

4. Repeat 1 and 2, but after standardizing each variable and replacing them with z-scores. Variables with higher variances could dominate the euclidean distances used in creating clusters. Did K change? What about the cluster memberships?

```{r}
set.seed(5)
#standardized the data 
df = all.us.city.crime.1970
df1 = all.us.city.crime.1970
#head(df,5)
dim(df)
for (i in 5:10) {
  df[,i] = (df1[,i] - mean(df1[,i])) / sd(df1[,i])
}
#df[,5] = (df1[,5] - mean(df1[,5])) / sd(df1[,5])
```

```{r}
#use WSS against K plot
## save WSS(K) for K=2,10)
WSS=NULL
for(K in 2:10){
 kms=kmeans(df[,5:10],centers=K,nstart=10)
 wssk=sum(kms$withinss)
 WSS=c(WSS,wssk)
}

plot(c(2:10),WSS,xlab="K",ylab="WSS(K)")
```

```{r}
#by maximizing the average silhouette
AS=c(1:9)
for(k in 2:10){
 km=kmeans(df[,5:10],centers=k,nstart=10)
 AS[k-1]=avsil(df[,5:10],km)
}
plot(c(2:10),AS)
```

We use K = 10
```{r}
set.seed(5)
cityclusters=kmeans(df[,5:10],centers=10)
city = all.us.city.crime.1970[,1]


city[cityclusters$cluster==1]
city[cityclusters$cluster==2]
city[cityclusters$cluster==3]
city[cityclusters$cluster==4]
city[cityclusters$cluster==5]
city[cityclusters$cluster==6]
city[cityclusters$cluster==7]
city[cityclusters$cluster==8]
city[cityclusters$cluster==9]
city[cityclusters$cluster==10]
```




5. Using these standardized data inspect the cluster means and use them to describe differences between clusters.

```{r}
set.seed(5)
km=kmeans(df[,5:10],centers=10,nstart=10)
km
```
So the 1st cluster means has the biggest murder rate
       2nd cluster means has the biggest rape rate
       3rd cluster means has the biggest assult rate
       4th cluster means has the second biggest burglary rate
       5th cluster means has the robbery murder rate
       
       cities above representing cities with serious crimes
       
       6th cluster means has the biggest car theft rate
       7th cluster means has decrease in all other crime but increase in car theft, representing cities with mild crime
       8th cluster means has decrease in all crime rates
       9th cluster means has mild increase in most crime rate
       10th cluster means has the lowest rape and murder rate, representing good cities.


